---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A common task in data engineering is to automate, schedule, and monitor multiple data processing pipelines. This is called **orchestration**. `maestro` is an R package that helps orchestrate data pipelines.

A fully realized `maestro` project involves the following components and actions:

1.  Collection of **pipelines** (R functions to be orchestrated, such as batch ETL jobs)

2.  **Orchestrator** - an R script or Quarto doc that orchestrates the pipelines and monitors them

3.  A process external to R to schedule the orchestrator (e.g., cron, Posit Connect).

### Project Setup

```{r}
library(maestro)
```

Create a `maestro` project in an existing project or a new project using `create_maestro()` or the New Project wizard in RStudio. This creates the orchestrator script and the folder of pipelines with one sample pipeline. Your project should look something like this:

```         
maestro_project
├── maestro_project.Rproj
├── orchestrator.R
└── pipelines
    ├── my_pipe.R
    └── another_pipe.R
```

## Pipelines

Pipelines are the jobs you want to automate, schedule, and monitor. For the most part, they're regular R functions with a special sprinkling of comments.

### Anatomy of a Pipeline

A pipeline is simply an R function with decorators called maestro tags. Maestro tags are special code comments used for communicating the scheduling and configuration of a pipeline to the orchestrator. Let's take a quick look at the sample `my_pipe.R`:

```{r}
#' my_pipe maestro pipeline
#'
#' @maestroFrequency 1 day
#' @maestroStartTime 2024-05-24
#' @maestroTz UTC
#' @maestroLogLevel INFO

my_pipe <- function() {

  # Pipeline code
}
```

`my_pipe` is a function with an empty body - so right now it won't do anything. The comments above are interpreted by `maestro` as "this function is scheduled to run every 1 day starting at 2024-05-24 (00:00:00) UTC time".

Note that you don't need to provide all these tags. A single maestro tag is enough to distinguish it as a pipeline. Pipelines missing tags will use consistent defaults (e.g., if `maestroFrequency` is missing the default is 1 day).

In most use cases, the actual code inside of `my_pipe` would be to run an ETL job (extract data from a source, transform it, and load it into a file system or database). In technical terms, it's the *side effect* of the code and not its return value that is important.

Here's a more realistic, albeit impractical, example:

```{r}
#' my_pipe maestro pipeline
#'
#' @maestroFrequency 1 day
#' @maestroStartTime 2024-05-24
#' @maestroTz UTC
#' @maestroLogLevel INFO

my_pipe <- function() {

  random_data <- data.frame(
    letters = sample(letters, 10),
    numbers = sample.int(10)
  )
  
  write.csv(random_data, file = tempfile())
}
```

### Adding New Pipelines

A project with a single pipeline is ok, but in `maestro` is more useful when you have multiple jobs to run. You can add more pipelines to your pipelines directory manually or use `create_pipeline()`:

```{r example-pipe, echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE}
create_pipeline(
  pipe_name = "another_pipeline",
  pipeline_dir = "pipelines",
  frequency = "1 hour",
  start_time = "2024-05-17 15:00:00",
  tz = "America/Halifax",
  log_level = "ERROR"
)
```

## Orchestrator

The orchestrator is the process that schedules and monitors the pipelines.

### Anatomy of the Orchestrator

The orchestrator can be an R script, Quarto/RMarkdown doc, but here we'll use a regular R script. Here is where you'll run `maestro` functions. The two main functions are `build_schedule()` and `run_schedule()`.

```{r orch, echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE}
library(maestro)

schedule_table <- build_schedule()

output <- run_schedule(
  schedule_table,
  orch_frequency = "1 hour"
)
```

Building the schedule gets `maestro` to look through the pipelines in the pipelines folder and creates a data.frame with schedule information. Then, you pass that data.frame to `run_schedule()` along with how often the orchestrator is supposed to run. It is important to tell `maestro` how often it'll be checking the pipelines using the `orch_frequency` parameter. Here, we're informing it that the orchestrator is running every 1 hour.

Importantly, it isn't `maestro`'s job to actually run it this often - it's *your* job to make sure it runs at that frequency (e.g., deploying it via cron or some cloud environment where code can be scheduled).[^1]

[^1]: The decision around how often to run the orchestrator depends on the frequencies of the pipelines in the project. The simplest guideline is to take the frequency of your most often recurring pipeline and split that in half. So for example, if my most often running pipeline runs every 1 day then my orchestrator should run every 12 hours.
